Results

### 2024-08-13 scripts/Results.Rmd
explore predicitons

  wide <- dplyr::mutate(
    x, 
    suitability = suitability, 
    .before = geometry) |>
    select(-f_suitability, -f_rankedp)

  # now we can join the class and patch metrics to each population 
  # which they were calculated for. 
  IDs <- tidyr::drop_na(wide, PchIDComb) |>
    dplyr::pull(PchIDComb)
  patch <- read.csv(sf::st_drop_geometry(x$f_patchMetrics[1]) ) |>
    dplyr::filter(id %in% IDs) |>
   # dplyr::rename(patch_metric = metric, patch_value = value) |>
    tidyr::pivot_wider(values_from = value, names_from = metric)
  class <- read.csv(sf::st_drop_geometry(x$f_classMetrics[1])) |>
    tidyr::pivot_wider(names_from = metric, values_from = value)

  ob <- cbind(patch, class)

  return(ob)

@@

scouting <- lapply(scout_split[1], extractR)
scouting <- dplyr::bind_rows(scouting)

scouting[[1]] |> 
    tidyr::pivot_wider(names_from = metric, values_from = value)

@@

```
```{r}
quants
```

@@

>>>>

  # now we can join the class and patch metrics to each population 
  # which they were calculated for. 
  IDs <- tidyr::drop_na(x, PchIDComb) |>
    dplyr::pull(PchIDComb)

  patch <- read.csv(sf::st_drop_geometry(x$f_patchMetrics[1]) ) |>
    dplyr::filter(id %in% IDs) |>
    tidyr::pivot_wider(values_from = value, names_from = metric)
  class <- read.csv(sf::st_drop_geometry(x$f_classMetrics[1])) |>
    tidyr::pivot_wider(names_from = metric, values_from = value)

  if(nrow(patch) == 0){
    patch <- data.frame(id = NA, cai = NA, enn = NA, frac = NA, para = NA )
  }

  lsm <- cbind(patch, class) |>
    dplyr::mutate(id = as.character(id))

  wide <- dplyr::mutate(
    x, 
    suitability = suitability, 
    .before = geometry) |>
    dplyr::select(-f_suitability, -f_rankedp, -f_patchMetrics, -f_classMetrics) |>
    dplyr::left_join(lsm, by = c('PchIDComb' = 'id' ))

  return(wide)

@@

scouting <- lapply(scout_split, extractR)
scouting <- dplyr::bind_rows(scouting)

```{r determine a null model for each species, eval = F}
#' return a large sample from a raster to plot the amount of suitable habitat 
#' 
#' sample many rasters to make a plot showing the cumulative amount of suitable habitat
#' @param x a data frame split by species including columns with paths to files. 
prcntSuitable <- function(x){

  r <- terra::rast(sf::st_drop_geometry(x$f_suitability[1]))
  Suitability = spatSample(r, size = 50000, na.rm = TRUE)
  q <- quantile(Suitability$predicted_suitability, probs = seq(0.01, 1, 0.01))

  sample <- 
    data.frame(
      Species = sf::st_drop_geometry(x$taxa[1]), 
      Percent = 1:100, 
      Records = as.numeric(q)
    )

  return(sample)

suitability_amounts <- lapply(scout_split, prcntSuitable)
saveRDS(suitability_amounts, file = '../data/processed/prcntSDMsurfaceSuitable.rds')
```

suitability_amounts <- readRDS('../data/processed/prcntSDMsurfaceSuitable.rds')
suitability_amounts <- bind_rows(suitability_amounts)

suitability_means <- suitability_amounts |>
  group_by(Percent) |>
  summarise(Mean = mean(Records))

par(pty="s")
plot(suitability_means$Mean, suitability_means$Percent, type = "l", col = "red", 
     ylab = "% of all Raster Cells", xlab = "Cumulative Sum of Records at each Predicted Suitability")

ggplot(suitability_amounts, aes(Percent, Records, color = Species, group = Species)) +
  geom_line() +
  scale_color_grey(guide = 'none') + 
  geom_line(data = suitability_means, 
            aes(Percent, Mean), inherit.aes = F, color = '#093824', lwd = 1) + 
  labs(
    y = "% of all Raster Cells",
    x = 'Cumulative sum of records at each quantile of Predicted Suitability') + 
  theme_classic() +
  theme(aspect.ratio = 1/1)

'#292E1E'
'#1F271B'
'#003E1F'
'#003E1F'
'#093824'
'#04471C'
'#1F2F16'
'#137547'
'#2A9134'

'#254441'
'#395B50'
'#73BA9B'
'#519872'

@@

```{r}
quants
```

@@

scout_long <- pivot_longer(scouting, cai:enn_mn)

ggplot(scout_long, aes(y = POPULATION_SIZE, x = value)) +
  geom_point() +
  facet_grid(. ~ name)
